The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Small changes throughout

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position, due
to relatively small-scale variations in the density of the gas and stellar
distribution. Converting positions and velocities of more distant stars from a
heliocentric to a Galactocentric coordinate system introduces a systematic
bias in the $z$ coordinate, and hence the actions computed for these stars.
The further from the solar neighborhood the target star, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the LSR.

@@

In Section~\ref{sec:discussion} we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses. In particular, we consider the impact on modelling
the distribution function of the Milky Way in Section~\ref{ssec:df}. We
summarize our main results and conclude in Section~\ref{sec:conclusion}.

@@

Now suppose this observer makes a measurement of the star's position and
velocity at the purple point or the green point (i.e. at two different orbital
phases). Then the purple and green curves correspond to the orbits that the
observer would compute for each point based on the potential model with the
offset midplane. In action space, this would correspond to a different value
of $J_z$ for the purple and green points. In this way, assuming the wrong
coordinate system induces a phase-dependence in the actions estimated for the
star, which in the correct potential (in this example, the one with the
correct midplane) should be phase-independent.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/cartoon.pdf}
\end{center}
\caption{Cartoon approximation showing the effect an error in the
determination of the coordinate midplane can have on orbit integration and
action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
vertical height. The top gray curve depicts an example ``true'' orbit
oscillating about the true midplane (horizontal solid gray line). Consider an
observer who erroneously assumes the midplane is located at the horizontal
dashed line. Suppose this observer measures the phase-space position of the
star at two different orbital phases (purple and green points). If the
observer were then to integrate the star's orbit using a model potential with
the erroneous midplane, they would obtain the purple and green curves for the
star's orbit, respectively. The actions estimated from these two  erroneous
orbits would subsequently differ, both from each other and from the actions
estimated for the true orbit (in the potential with the correct midplane).
Hence an incorrect midplane in the potential model assumed will induce
phase-dependence in the actions estimated for a given star in that potential.}
\label{fig:cartoon}
\end{figure*}

\subsection{Epicyclic Approximation} \label{ssec:epi_action}
Before turning to numerical methods, we first write down analytically the
systematic error in the actions induced from either a midplane offset or a
Galactocentric radius offset. We use the epicyclic approximation, which
assumes that the motion in the $z$ and $R$ components of the orbit are
decoupled and follow simple harmonic motion about a circular and planar
``guiding orbit'' \citep[][Section 3.2 and references
therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
``guiding radius'') as $R_g$. This approximation is an excellent description
of thin disk-like orbits and a good description of thick disk-like orbits in
an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
spiral arms). We also make the assumption of a perfectly flat circular
velocity curve with velocity $v_c$ \citep[decent near the Solar circle,
e.g.][]{2017MNRAS.465...76M}.

@@

In this case, we have that for $J_{\phi}$
\citep[][Section~3.5.3b]{2008gady.book.....B}:
\beq\label{eq:Jphi_epi}
J_{\phi} = R_g v_c\text{,}
\eeq
and for $J_R$ and $J_z$:
\beq\label{eq:JR_Jz_epi}
\begin{split}
J_R &= \frac{E_R}{\kappa} \\
J_z &= \frac{E_z}{\nu} \text{,}
\end{split}
\eeq
where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
coordinates, respectively. Therefore,
\beq\label{eq:JR_Jz_epi_energy}
\begin{split}
J_R &= \frac{v_R^2 + \kappa^2 (R-R_g)^2}{2\kappa} \\
J_z &= \frac{v_z^2 + \nu^2 z^2}{2\nu}\text{.}
\end{split}
\eeq
We can simplify this and write
\beq\label{eq:JR_Jz_epi_final}
\begin{split}
J_R &= \frac{\kappa A_R^2}{2} = \frac{v_{R,\text{max}}^2}{2\kappa} \\
J_z &= \frac{\nu A_z^2}{2} = \frac{v_{z,\text{max}}^2}{2\nu}\text{,}
\end{split}
\eeq
where the last equality comes from the fact that $v_{R,\text{max}} = \kappa
A_R$ and similarly for $z$.

@@

In the remainder of this section, we compare our analytic estimates of the
effect on actions of a midplane offset against numerical calculations. A
numerical evaluation of the effect of velocity offsets on actions is deferred
to future work, as we discuss in Section~\ref{ssec:lsr_var}.

\subsection{Numerical Methods} \label{ssec:action_comp}
We now quantify the argument made in Section~\ref{ssec:cartoon} using
numerical computations of the actions for a range of orbits in a model
Galactic potential. We compute actions as in
\citet{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
orbit integrations and conversion to action space
\citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions we use the
torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and
adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital
time-series starting from a phase-space position $(x, v)$ and integrated in a
potential $\Phi$. We use the default \texttt{MWPotential} as our potential,
which is based on the Milky Way potential available in \texttt{galpy}
\citep{2015ApJS..216...29B}. This potential includes a Hernquist bulge and
nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
calculations depend on the value of the LSR in this toy potential (though this
is important when using real data, since the conversion from heliocentric to
Galactocentric coordinates depends on the LSR). In this potential, we have
that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ on the Solar
circle.

@@

We begin by considering the thick-disk orbit. Suppose an immortal observer who
can measure the thick-disk star's phase-space position at many different times
(and hence different orbital phases), but used a coordinate system in which
the midplane is systematically offset in height by $100\,\pc$ from its actual
location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
in the orbit. This corresponds to the immortal observer physically located at
e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
offset coordinate systems depend on the time (i.e. orbital phase) at which the
star's phase-space position is observed. To the immortal observer this would
appear as time-dependence in the actions, which appear to oscillate around
their value using the correct coordinate system even though in this example
the observer is using the correctly constructed best-fit static, axisymmetric
potential. The relative size of the phase variation in each action depends on
the direction of the systematic offset as well as the true values of the
actions (i.e. the type of orbit). In reality we will have one measurement of
the phase-space position to work with, in which case the determination of the
orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
therefore quote percentile ranges for the possible values computed for each
action as a proxy for the effect of these systematic errors in the coordinate
system.

@@

For a systematic offset in $x$ (or Galactic center distance), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
despite only a $1.2\%$ error in the distance to the Galactic center.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_one_orbit.pdf}
\end{center}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
integrate the orbit according to the procedure laid out in
Section~\ref{ssec:action_comp}, and which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$
value (upper panels) or the $x$ value (lower panels) of each
position in the orbit, corresponding to an erroneous observer assuming a
midplane (upper) or solar radius (lower) that is off by
$100\,\pc$. We then allow an immortal observer to observe the orbit over
$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
and report the values of the actions. The computation of $J_{\phi}$ is
pristine to errors in $z$, with only numerical artifacts remaining. Only small
errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
being within $\sim8\%$ of the true $J_R$. As expected, large errors are
induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_Jz_zerr_hist.pdf}
\end{center}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the error distribution.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel) assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper
right panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
is plotted as a vertical dashed line. Here we see that the errors in $J_z$
induced by a systematic offset in $z$ are non-Gaussian and bimodal.
Furthermore, neither of the modes lie at zero. In other words, neither of the
modes of the distribution of observed $J_z$ values are centered on the true
$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
thin-disk orbit (bottom panel), we see that, in addition to the prior
complications, the error distribution is not even centered on null. This comes
about when the midplane error is $\gtrsim z_{\text{max}}$, where
$z_{\text{max}}$ is the maximum height of the orbit (equivalent to $A_z$ in
Section~\ref{ssec:epi_action}).

@@

In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
include the estimation of the action errors derived under the epicyclic
approximation from Equation~\eqref{eq:Ji_err_mosttime}. This Equation is
relevant since during most of the orbit the star will be at maximum radial and
vertical amplitude and we do not consider velocity offsets. Note that we
consider an error in the $x$-coordinate $\Delta x$, which is not quite the
same as $\Delta R$. For observations of stars close to us, we have that
$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
consider observations of the star throughout its entire orbit. This introduces
a factor of $2/\pi$ when converting from $\Delta R$ to $\Delta x$, which we
derive in Appendix~\ref{app:deltax}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_many_orbits_Jz_fun.pdf}
\end{center}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one
half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
$J_z$ values over the course of the orbit. There are large errors for
thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
offset of $10\,\pc$. }
\label{fig:dJz_fun_Jz}
\end{figure}

To further understand the effect of the midplane error, we also plot the
fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
100\,\pc$ (orange, cyan, and green, respectively) in
Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
are suppressed, but the induced error can still be large depending on how
great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
offset. We find that the epicyclic approximation matches the numerical
estimate quite well.

@@

Importantly, the location of \sgra{} may not be equivalent to
the location of the ``dynamical Galactic center:'' the point in
three-dimensional space about which the stars in the solar neighborhood are
orbiting. This assumption, although sensible and frequently made, has not yet been
justified. It is tested in any construction of a  dynamical model where the
distance from the Sun to the Galactic center, $R_0$, is a free parameter. For
example, \citet{2015ApJ...803...80K} measured $R_0$ while modelling the
dynamics of the stream Palomar 5. Many other dynamical measurements of $R_0$
have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia}
results) but none have yet achieved a precision comparable to that of the
distance to \sgra{}.

\subsection{Galactic Orientation}
Second, one must define the angular orientation of the Galaxy. This was
defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
defining the coordinates of the Galactic center in B1950 coordinates as
(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00).
These two quantities, together with $R_0$, define the orientation of the
Galactic plane. However, there is growing evidence that the stellar midplane
is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
\citep{2019ApJ...871..145A}.

@@

\subsection{Description of FIRE Simulations} \label{ssec:cosmozoom}
The FIRE cosmological hydrodynamic simulations
\citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
of a small group of galaxies at high resolution in a full cosmological
context.  Feedback from supernovae, stellar winds, and radiation from massive
stars is implemented at the scale of star forming regions following stellar
population synthesis models, generating galactic winds self-consistently
\citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while reproducing many
observed galaxy properties, including stellar masses, star formation
histories, metallicities, and morphologies and kinematics of thin and thick
disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}. For this work,
we focus on the three Milky Way-mass zoom-ins considered in
\citet{2018arXiv180610564S}, which were simulated as part of the
\textit{Latte} suite and show broad agreement of many of their global
properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
2018MNRAS.481.4133G} The $\z = 0$ snapshots\footnote{In this work, to avoid
confusion with the vertical height $z$, we refer to cosmological redshift as
$\z$.} of these three simulations, named \mi{}, \mf{}, and \mm{} for
shorthand, are publicly available alongside associated mock \textit{Gaia} DR2
catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}

>>>>

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, their local midplane location
will not agree with our local midplane extrapolated onto their position due
to relatively small-scale variations in the density of the gas and stellar
distribution. Converting positions and velocities of more distant stars from a
heliocentric to a Galactocentric coordinate system introduces a systematic
bias in the $z$ coordinate, and hence the actions computed for these stars.
The further from the solar neighborhood the target star, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the LSR.

@@

In Section~\ref{sec:discussion}, we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses. We also estimate the expected midplane variations of the Milky Way based on the observed velocity variations. We
summarize our main results and conclude in Section~\ref{sec:conclusion}.

@@

Now suppose this observer makes a measurement of the star's position and
velocity at the teal point or the orange point (i.e. at two different orbital
phases). Then the teal and orange curves correspond to the orbits that the
observer would compute for each point based on the potential model with the
offset midplane. In action space, this would correspond to a different value
of $J_z$ for the teal and orange points. In this way, assuming the wrong
coordinate system induces a phase-dependence in the actions estimated for the
star, which in the correct potential (in this example, the one with the
correct midplane) should be phase-independent.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/cartoon.pdf}
\end{center}
\caption{Cartoon approximation showing the effect an error in the
determination of the coordinate midplane can have on orbit integration and
action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
vertical height. The top gray curve depicts an example ``true'' orbit
oscillating about the true midplane (horizontal solid gray line). Consider an
observer who erroneously assumes the midplane is located at the horizontal
dashed line. Suppose this observer measures the phase-space position of the
star at two different orbital phases (teal and orange points). If the
observer were then to integrate the star's orbit using a model potential with
the erroneous midplane, they would obtain the teal and orange curves for the
star's orbit, respectively. The actions estimated from these two  erroneous
orbits would subsequently differ, both from each other and from the actions
estimated for the true orbit (in the potential with the correct midplane).
Hence an incorrect midplane in the potential model assumed will induce
phase-dependence in the actions estimated for a given star in that potential.}
\label{fig:cartoon}
\end{figure*}

\subsection{Epicyclic Approximation} \label{ssec:epi_action}
Before turning to numerical methods, we first write down analytically the
systematic error in the actions induced from either a midplane offset or a
Galactocentric radius offset. We use the epicyclic approximation, which
assumes that the motion in the $z$ and $R$ components of the orbit are
decoupled and follow simple harmonic motion about a circular and planar
``guiding orbit'' \citep[][Section~3.2 and references
therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
``guiding radius'') as $R_g$. This approximation is an excellent description
of thin disk-like orbits and a good description of thick disk-like orbits in
an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
spiral arms). We also make the assumption of a perfectly flat circular
velocity curve with velocity $v_c$, a good approximation near the Solar circle \citep[e.g.][]{2017MNRAS.465...76M}.

@@

In this case, we have that for $J_{\phi}$
\citep[][Section~3.5.3b]{2008gady.book.....B}:
\beq\label{eq:Jphi_epi}
J_{\phi} = R_g v_c\text{,}
\eeq
and for $J_R$ and $J_z$:
\beq\label{eq:JR_Jz_epi}
\begin{split}
J_R &= \frac{E_R}{\kappa} \\
J_z &= \frac{E_z}{\nu} \text{,}
\end{split}
\eeq
where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
coordinates, respectively. Therefore,
\beq\label{eq:JR_Jz_epi_energy}
\begin{split}
J_R &= \frac{v_R^2 + \kappa^2 (R-R_g)^2}{2\kappa} \\
J_z &= \frac{v_z^2 + \nu^2 z^2}{2\nu}\text{.}
\end{split}
\eeq
Using Equations~\eqref{eq:orbits_epi} and \eqref{eq:orbits_vel_epi}, we can
simplify this and write
\beq\label{eq:JR_Jz_epi_final}
\begin{split}
J_R &= \frac{\kappa A_R^2}{2} = \frac{v_{R,\text{max}}^2}{2\kappa} \\
J_z &= \frac{\nu A_z^2}{2} = \frac{v_{z,\text{max}}^2}{2\nu}\text{,}
\end{split}
\eeq
where the last equality in each line comes from the fact that $v_{R,\text{max}} = \kappa
A_R$ and $v_{z,\text{max}} = \nu A_z$.

@@

In the remainder of this section, we compare our analytic estimates of the
effect of a midplane offset on actions against numerical calculations. A
numerical evaluation of the effect of velocity offsets on actions is deferred
to future work, as we discuss in Section~\ref{ssec:lsr_var}.

\subsection{Numerical Methods} \label{ssec:action_comp}
We now quantify the argument made in Section~\ref{ssec:cartoon} using
numerical computations of the actions for a range of orbits in a model
Galactic potential. We compute actions as in
\citet{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
orbit integrations and conversion to action space
\citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions we use the
torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and
adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital
time-series starting from a phase-space position $(x, v)$ and integrated in a
potential $\Phi$. We use the default \texttt{MWPotential} as our potential,
which is based on the Milky Way potential available in \texttt{galpy}
\citep{2015ApJS..216...29B}. This potential includes a Hernquist bulge and
nucleus \citep{1990ApJ...356..359H}, a Miyamoto--Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
calculations depend on the value of the LSR in this toy potential (though this
is important when using real data, since the conversion from heliocentric to
Galactocentric coordinates depends on the LSR). In this potential, we have
that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ at the Solar
circle.

@@

We begin with the thick-disk orbit. Consider an immortal observer who
can measure the thick-disk star's phase-space position at many different times
(and hence different orbital phases). However, the immortal observer uses a coordinate system in which
the midplane is systematically offset in height by $100\,\pc$ from its actual
location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
in the orbit. This corresponds to the immortal observer physically located at
e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
offset coordinate systems depend on the time (i.e. orbital phase) at which the
star's phase-space position is observed. To the immortal observer this would
appear as time-dependence in the actions, which appear to oscillate around
their value using the correct coordinate system even though in this example
the observer is using the correctly constructed best-fit static, axisymmetric
potential. The relative size of the phase variation in each action depends on
the direction of the systematic offset as well as the true values of the
actions (i.e. the type of orbit). In reality we will have one measurement of
the phase-space position to work with, in which case the determination of the
orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
therefore quote percentile ranges for the possible values computed for each
action as a proxy for the effect of these systematic errors in the coordinate
system.

@@

For a systematic offset in $x$ (or distance to the Galactic center), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
despite only a $1.2\%$ error in the distance to the Galactic center.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/schmactions_one_orbit.pdf}
\end{center}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
integrate the orbit according to the procedure laid out in
Section~\ref{ssec:action_comp}, and which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$
value (upper panels) or the $x$ value (lower panels) of each
position in the orbit, corresponding to an erroneous observer assuming a
midplane (upper) or solar radius (lower) that is off by
$100\,\pc$. We then allow an immortal observer to observe the orbit over
$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
and report the values of the actions. The computation of $J_{\phi}$ is
pristine to errors in $z$, with only numerical artifacts remaining. Only small
errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
being within $\sim8\%$ of the true $J_R$. As expected, large errors are
induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of $\sim21\%$, $\sim3\%$, and $\sim3\%$.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_Jz_zerr_hist.pdf}
\end{center}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the distribution of observed $J_z$ values.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel) assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper
right panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
is plotted as a vertical dashed line. Here we see that the errors in $J_z$
induced by a systematic offset in $z$ are non-Gaussian and bimodal.
Furthermore, neither of the modes lie at zero. In other words, neither of the
modes of the distribution of observed $J_z$ values are centered on the true
$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
thin-disk orbit (bottom panel), we see that, in addition to the prior
complications, the error distribution is not even centered on null. This comes
about when the midplane error is $\gtrsim z_{\text{max}}$, where
$z_{\text{max}}$ is the maximum height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
Section~\ref{ssec:epi_action}).

@@

In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
include the estimation of the action errors derived under the epicyclic
approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
relevant since during most of the orbit the star will be at maximum radial and
vertical amplitude and we do not consider velocity offsets. Note that we
consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
same as $\Delta R$. For observations of stars close to us, we have that
$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
consider observations of the star throughout its entire orbit. This introduces
a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
derive in Appendix~\ref{app:deltax}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_many_orbits_Jz_fun.pdf}
\end{center}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is one
half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
$J_z$ values over the course of the orbit. There are large errors for
thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
offset of $10\,\pc$. }
\label{fig:dJz_fun_Jz}
\end{figure}

To further understand the effect of the midplane error, we also plot the
fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
100\,\pc$ (orange, teal, and green, respectively) in
Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
are suppressed, but the induced error can still be large depending on how
great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
offset. We find that the epicyclic approximation matches the numerical
estimate quite well.

@@

Importantly, the location of \sgra{} may not be equivalent to
the location of the ``dynamical Galactic center,'' i.e., the point in
three-dimensional space about which the stars in the solar neighborhood are
orbiting. This assumption, although sensible and frequently made, has not yet been
justified. It is tested in any construction of a  dynamical model where the
distance from the Sun to the Galactic center, $R_0$, is a free parameter. For
example, \citet{2015ApJ...803...80K} measured $R_0$ while modelling the
dynamics of the stream Palomar 5. Many other dynamical measurements of $R_0$
have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia}
results) but none have yet achieved a precision comparable to that of the
distance to \sgra{}.

\subsection{Galactic Orientation}
Second, one must define the angular orientation of the Galaxy. This was
defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
defining the coordinates of the Galactic center in B1950 coordinates as
(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
These two quantities, together with $R_0$, define the orientation of the
Galactic plane. However, there is growing evidence that the stellar midplane
is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
\citep{2019ApJ...871..145A}.

@@

\subsection{Description of FIRE Simulations} \label{ssec:cosmozoom}
The FIRE cosmological hydrodynamic simulations
\citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
of a small group of galaxies at high resolution in a full cosmological
context.  Feedback from supernovae, stellar winds, and radiation from massive
stars is implemented at the scale of star forming regions following stellar
population synthesis models, generating galactic winds self-consistently
\citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while reproducing many
observed galaxy properties, including stellar masses, star formation
histories, metallicities, and morphologies and kinematics of thin and thick
disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}. For this work,
we focus on the three Milky Way-mass zoom-ins considered in
\citet{2018arXiv180610564S}, which were simulated as part of the
\textit{Latte} suite and show broad agreement of many of their global
properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
2018MNRAS.481.4133G}. The $\z = 0$ snapshots\footnote{In this work, to avoid
confusion with the vertical height $z$, we refer to cosmological redshift as
$\z$.} of these three simulations, named \mi{}, \mf{}, and \mm{} for
shorthand, are publicly available alongside associated mock \textit{Gaia} DR2
catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$, with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with galactocentric cylindrical radii within $0.5\,\kpc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and $10.3\,\pc$ for \mi{},
\mf{}, and \mm{}, respectively. These values are summarized in
Table~\ref{tab:scale_height}, along with measurements of the stellar and gas
disk scale heights.
